---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# astrocoords

<!-- badges: start -->

[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)

<!-- badges: end -->

`astrocoords` is an R package for working with astronomical coordinates and matching catalogs.

It stores sky coordinates as a vector type instead of separate numeric columns, making coordinate transformations, angular separations, catalog matching, and spatial joins more natural in R.


It provides:

- sky coordinate objects (`sky_coord`) with ICRS and Galactic frames

- angular separation and frame transformations

- catalog matching (`coord_match()`, `coord_nearest()`)

- spatial joins for data frames (`coord_left_join()`, `coord_nearest_join()`, etc.)

- sexagesimal conversion helpers

- Julian Date / Modified Julian Date conversion

Core coordinate and time calculations are based on the [ERFA](https://github.com/liberfa/erfa) library.

## Installation

You can install the development version of astrocoords from [GitHub](https://github.com/uskovgs/astrocoords) with:

```r
# install.packages("pak")
pak::pak("uskovgs/astrocoords")
```

## Quick examples

Create coordinates in the ICRS frame:

```{r}
library(astrocoords)

x <- ra_dec(c(10, 20), c(30, 40))
x
```

Transform them to Galactic coordinates:

```{r}
transform_to(x, galactic())
```

Measure angular separation in arcseconds:

```{r}
separation(ra_dec(10, 20), ra_dec(11, 21))
```

Find nearest matches between two coordinate vectors:

```{r}
coord_nearest(
  ra_dec(c(10, 20), c(0, 0)),
  ra_dec(c(10.001, 19.999), c(0, 0)),
  unit = "arcsec",
  method = "kdtree"
)
```

Use a spatial join between two data frames:

```{r}
df1 <- data.frame(
  id = 1:2,
  coord = ra_dec(c(10, 20), c(30, 40))
)

df2 <- data.frame(
  name = c("a", "b"),
  coord = ra_dec(c(10.0002, 20.0003), c(30.0001, 39.9998))
)

df1 |>
  coord_left_join(df2, max_sep = 5, unit = "arcsec")
```

Parse a compact coordinate string:

```{r}
sc <- parse_coord("J230631.0+155633")
sc
```

```{r}
iau_name(sc, prefix = "SRGA J", ra_digits = 3, dec_digits = 2)
```

Work with Julian Date:

```{r}
jd <- datetime_to_jd(as.POSIXct("2026-03-08 12:34:56", tz = "UTC"))
print(jd, digits=11)

jd_to_datetime(jd, tz = "UTC")
```


## Learn more

- Website: <https://uskovgs.github.io/astrocoords/>
- Reference: <https://uskovgs.github.io/astrocoords/reference/>
- Articles: <https://uskovgs.github.io/astrocoords/articles/>