---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# astrocoords

<!-- badges: start -->

[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)

<!-- badges: end -->

`astrocoords` is a small R package for basic astronomical coordinates and time conversion.

Core coordinate and time calculations in `astrocoords` are based on the
[ERFA](https://github.com/liberfa/erfa) library.

It currently supports:

- ICRS and Galactic sky coordinates
- angular separation on the sphere
- simple frame transformation between ICRS and Galactic
- Julian Date / Modified Julian Date conversion

## Installation

You can install the development version of astrocoords from [GitHub](https://github.com/uskovgs/astrocoords) with:

```r
# install.packages("pak")
pak::pak("uskovgs/astrocoords")
```

## Quick examples

Create coordinates in the ICRS frame:

```{r}
library(astrocoords)

x <- ra_dec(c(10, 20), c(30, 40))
x
```

Transform them to Galactic coordinates:

```{r}
transform_to(x, galactic())
```

Measure angular separation in arcseconds:

```{r}
separation(ra_dec(10, 20), ra_dec(11, 21))
```

Parse a compact coordinate string:

```{r}
parse_coord("J230631.0+155633")
```

Work with Julian Date:

```{r}
jd <- jd_fromdate(as.POSIXct("2026-03-08 12:34:56", tz = "UTC"))
jd

jd2greg(jd, tz = "UTC")
```
