---
title: "Get started"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{get_started}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(astrocoords)
```

This vignette shows the basic ideas of the package.

You can create sky coordinates in the ICRS frame or in the Galactic frame.

```{r}
ra_dec(10, 20)
gal_coord(120, -30)
```

The general constructor is `sky_coord()`. You choose the frame with `icrs()` or `galactic()`.

```{r}
sky_coord(c(10, 20), c(30, 40), frame = icrs())
sky_coord(c(120, 121), c(-30, -31), frame = galactic())
```

You can measure angular separation between two points. The result is in arcseconds.

```{r}
x1 <- ra_dec(10, 20)
x2 <- ra_dec(11, 21)

separation(x1, x2)
```

You can also transform coordinates between ICRS and Galactic.

```{r}
x <- ra_dec(c(10, 20), c(30, 40))
g <- transform_to(x, galactic())
g

transform_to(g, icrs())
```

Accessors depend on the frame. Use `ra()` and `dec()` for ICRS, and `l()` and `b()` for Galactic.

```{r}
x <- ra_dec(10, 20)
ra(x)
dec(x)

g <- gal_coord(120, -30)
l(g)
b(g)
```

You can also parse compact ICRS strings.

```{r}
parse_coord("J230631.0+155633")
```

You can build a simple IAU-style source name from ICRS coordinates.

```{r}
parse_coord("J230631.0+155633") |>
  iau_name(prefix = "SRGA J")
```
