---
title: "Catalog matching joins"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{coord_matching}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(astrocoords)
library(ggplot2)
theme_set(theme_light())
```

## TL;DR

- `coord_inner_join()` keeps only rows from `x` that have matches in `y` within `max_sep`.
- `coord_left_join()` keeps all rows from `x`; unmatched rows get `NA` from `y`.
- `coord_right_join()` keeps all rows from `y`; unmatched rows get `NA` from `x`.
- `coord_full_join()` keeps all rows from both tables.
- `coord_nearest_join()` returns one nearest match in `y` for each row in `x`.
- `multiple = "all"` keeps all matches inside the radius, while `multiple = "closest"` keeps only one match per row in `x`.




# Understanding joins

```{r}
set.seed(1)
N <- 10
df1 <- data.frame(
  ra = runif(N), 
  dec = runif(N),
  cat = "first"
)
df2 <- data.frame(
  ra = runif(N), 
  dec = runif(N),
  cat = "second"
)

df12 <- do.call(rbind, list(df1, df2))
df1$sc <- ra_dec(df1$ra, df1$dec)
df2$sc <- ra_dec(df2$ra, df2$dec)

df12$sc <- c(df1$sc, df2$sc)

p1 <- df12 |>
  ggplot(aes(ra, dec, color = cat)) +
  geom_point() +
  coord_fixed(ratio = 1)
p1
```

## Inner join within radius

```{r}
df12_intersect <- df1 |>
  coord_inner_join(df2, max_sep = 5, unit = "arcmin")

p1 +
  geom_segment(
    data = df12_intersect,
    aes(x = ra.x, xend = ra.y, y = dec.x, yend = dec.y),
    color = "black"
  ) +
  ggforce::geom_circle(
    data = df12_intersect,
    aes(x0 = ra.x, y0 = dec.x, r= 5/60),
    fill = "grey50",
    alpha = 0.2,
    inherit.aes = FALSE
  ) +
    geom_label(
      data = df12_intersect,
      aes(ra.x, dec.x + 0.1, label = paste(round(sep, 1), "arcmin")),
      inherit.aes = FALSE,      
    )
```


## Nearest-neighbor join


```{r}
df12_nearest <- df1 |>
  coord_nearest_join(df2, unit = "arcmin") 

p1 +
  geom_segment(
    data = df12_nearest,
    aes(x = ra.x, xend = ra.y, y = dec.x, yend = dec.y),
    color = "black"
  ) +
  ggforce::geom_circle(
    data = df12_nearest,
    aes(x0 = ra.x, y0 = dec.x, r = sep/60),
    fill = "grey50",
    alpha = 0.2,
    inherit.aes = FALSE
  ) +
    geom_label(
      data = df12_nearest,
      aes(ra.x, dec.x + 0.1, label = paste(round(sep, 1), "arcmin")),
      inherit.aes = FALSE
    )
```


## Left vs right vs full join

```{r}
df1 |>
  coord_left_join(df2, max_sep = 5, unit = "arcmin")
```


```{r}
df1 |>
  coord_right_join(df2, max_sep = 5, unit = "arcmin")
```

```{r}
df1 |>
  coord_full_join(df2, max_sep = 5, unit = "arcmin")
```

# Multiple matches: all vs closest


```{r}
set.seed(1)

N <- 13
df1 <- data.frame(
  ra = runif(N), 
  dec = runif(N),
  cat = "first"
)
df2 <- data.frame(
  ra = runif(N), 
  dec = runif(N),
  cat = "second"
)

df12 <- do.call(rbind, list(df1, df2))
df1$sc <- ra_dec(df1$ra, df1$dec)
df2$sc <- ra_dec(df2$ra, df2$dec)

df12$sc <- c(df1$sc, df2$sc)

p1 <- df12 |>
  ggplot(aes(ra, dec, color = cat)) +
  geom_point() +
  coord_fixed(ratio = 1)
p1
```

## `multiple="all"`
```{r}
MAX_SEP <- 8
df12_intersect <- df1 |>
  coord_inner_join(df2, max_sep = MAX_SEP, unit = "arcmin")

p1 +
  geom_segment(
    data = df12_intersect,
    aes(x = ra.x, xend = ra.y, y = dec.x, yend = dec.y),
    color = "black"
  ) +
  ggforce::geom_circle(
    data = df12_intersect,
    aes(x0 = ra.x, y0 = dec.x, r = MAX_SEP/60),
    fill = "grey50",
    alpha = 0.2,
    inherit.aes = FALSE
  )
```

## `multiple="closest"`

```{r}
df12_intersect <- df1 |>
  coord_inner_join(df2, max_sep = MAX_SEP, unit = "arcmin", multiple = "closest")

p1 +
  geom_segment(
    data = df12_intersect,
    aes(x = ra.x, xend = ra.y, y = dec.x, yend = dec.y),
    color = "black"
  ) +
  ggforce::geom_circle(
    data = df12_intersect,
    aes(x0 = ra.x, y0 = dec.x, r= MAX_SEP/60),
    fill = "grey50",
    alpha = 0.2,
    inherit.aes = FALSE
  )
```
